---
title: "analysis"
format: html
---

```{r}
library(tidyverse)
library(here)
library(glue)
library(jsonlite)
library(patchwork)
library(lme4)
library(lmerTest)
library(brms)
library(udpipe)

source(here("analysis_scripts", "logprobs.R"))
source(here("analysis_scripts", "plot_helpers.R"))

CONTEXT_LOC <- here("context_prep")
OUTPUT_LOC <- here("data", "logprobs")
```

# Load data

Exp 3 data
```{r}
human_data <- read_csv(url("https://raw.githubusercontent.com/vboyce/tg-matcher/refs/heads/main/data/tgmatcheryoked-trials.csv"))

human_selections <- human_data |>
  filter(!is.na(correct_tangram)) |>
  filter(type == "selection") |>
  select(
    workerid, correct, correct_tangram, condition,
    gameId, selected, trial_index, type, orig_trialNum, orig_repNum
  ) |>
  mutate(workerid = as.factor(workerid),
         matcher_trialNum = (trial_index - 3) %/% 3,
         matcher_repNum = matcher_trialNum %/% 12,
         # somehow two participants were assigned to 3157/139 -- but each set looks complete?
         workerid = ifelse(workerid == "3157" & condition == "yoked", "3157a", workerid),
         workerid = ifelse(workerid == "139" & gameId == "Z2WAkYpWiXwuGcdME", "139a", workerid)) |> 
  filter(workerid != "141", workerid != "35") |> # exclude two participants who didn't finish
  select(-trial_index)
```

Exp 5 data
```{r}
human_data_new <- read_csv(url("https://raw.githubusercontent.com/vboyce/tg-matcher/refs/heads/main/data/tgmatcher5.csv"))

human_selections_new <- human_data_new |>
  filter(!is.na(correct_tangram),
         type == "selection") |>
  select(
    workerid, correct, correct_tangram, condition,
    gameId, selected, trial_index, type, orig_trialNum, orig_repNum
  ) |>
  mutate(workerid = as.factor(workerid),
         matcher_trialNum = (trial_index - 3) %/% 3,
         matcher_repNum = matcher_trialNum %/% 12,
         workerid = case_when( # some error in id assignment so that two prolific id got the same assigment
           workerid == 4969 & row_number() > 5000 ~ "4969a",
           workerid == 5003 & row_number() > 10000 ~ "5003a",
           T ~ as.character(workerid)
         )) |> 
  filter(workerid != "4957") |> # exclude participant who didn't finish
  select(-trial_index)
```


Original data
```{r}
choose <- c(
  "Z2WAkYpWiXwuGcdME", "GWCD2NGoiA2n5Wh2i", "zviCQXNM4xWoYTmoY", "MEJ8y7jRW4947MPrH",
  "x8BXJBHh3RxchPoHv", "rSMFkjhskBteiyQMB", "LzNv7CD7gRuxpk7cQ", "MDLyuv29jevaGLSeT",
  "BTbGhXZvjdSFubTBg", "22dyGMRgestp8u5Lc"
)

original_chat <- read_csv(url("https://raw.githubusercontent.com/vboyce/tg-matcher/refs/heads/main/expt_prep_code/combined_chat.csv")) |> 
  filter(gameId %in% choose)
```

## Reshape human data
```{r}
human_orig <- original_chat |>
  mutate(
    role = case_when(
      role == "speaker" ~ "describer",
      role == "listener" ~ "matcher"
    ),
    playerCount = coalesce(activePlayerCount, numPlayers),
    accuracy = realCorrect / (playerCount - 1)
  ) |>
  mutate(
    matcher_trialNum = trialNum,
    matcher_repNum = repNum,
    type = "human_original",
    condition = "yoked"
  ) |>
  select(gameId, condition, orig_trialNum = trialNum, orig_repNum = repNum, 
         matcher_trialNum, matcher_repNum, target = tangram, accuracy, type) |>
  distinct()
```

```{r}
yoked_human <- human_selections |> 
  filter(condition == "yoked") |>
  group_by(gameId, condition, orig_trialNum, orig_repNum, 
           matcher_trialNum, matcher_repNum, target = correct_tangram) |>
  summarise(accuracy = mean(correct),
            .groups = "drop") |> 
  mutate(type = "human_naive")

shuffled_human <- human_selections |> 
  filter(condition == "shuffled") |>
  group_by(gameId, condition, orig_trialNum, orig_repNum, 
           matcher_trialNum, matcher_repNum, target = correct_tangram) |>
  summarise(accuracy = mean(correct),
            .groups = "drop") |> 
  mutate(type = "human_naive")
```

```{r}
backward_human <- human_selections_new |> 
  filter(condition == "backwards") |>
  group_by(gameId, condition, orig_trialNum, orig_repNum, 
           matcher_trialNum, matcher_repNum, target = correct_tangram) |>
  summarise(accuracy = mean(correct),
            .groups = "drop") |> 
  mutate(type = "human_naive",
         condition = "backward")

random_human <- human_selections_new |> 
  filter(condition == "random") |>
  group_by(gameId, condition, orig_trialNum, orig_repNum, 
           matcher_trialNum, matcher_repNum, target = correct_tangram) |>
  summarise(accuracy = mean(correct),
            .groups = "drop") |> 
  mutate(type = "human_naive")
```


## Everything
```{r}
logprob_files <- list.files(here("data", "logprobs"), recursive = TRUE)

full_logprobs <- map(logprob_files, load_logprobs) |> list_rbind()

logprobs_base <- full_logprobs |> 
  filter_logprobs()
```

Fix lengths
```{r}
human_orig_fixed <- human_orig |> 
  cbind(logprobs_base |> 
          filter(type == "model_gemma",
                 feedback == "full",
                 condition == "yoked") |> 
          select(message_length, context_length)) |> 
  mutate(runId = "1")
yoked_human_fixed <- yoked_human |> 
  cbind(logprobs_base |> 
          filter(type == "model_gemma",
                 feedback == "full",
                 condition == "yoked") |> 
          select(message_length, context_length)) |> 
  mutate(runId = "1")
shuffled_human_fixed <- human_selections |> 
  filter(condition == "shuffled") |> 
  cbind(logprobs_base |> 
          filter(type == "model_gemma",
                 feedback == "full",
                 condition == "shuffled") |> 
          select(message_length, context_length)) |> 
  mutate(runId = as.character(workerid)) |>
  group_by(gameId, runId, condition, orig_trialNum, orig_repNum, 
           matcher_trialNum, matcher_repNum, target = correct_tangram) |>
  summarise(accuracy = mean(correct),
            message_length = mean(message_length),
            context_length = mean(context_length),
            .groups = "drop") |> 
  mutate(type = "human_naive")
```

```{r}
backward_human_fixed <- backward_human |> 
  cbind(logprobs_base |> 
          filter(type == "model_gemma",
                 feedback == "full",
                 condition == "backward") |> 
          select(message_length, context_length)) |> 
  mutate(runId = "1")
random_msgs <- read_csv(here("context_prep", "full_feedback", "random.csv")) |> 
  mutate(message = map(message, fromJSON),
         message_history = message_history |>
           replace_na("[]") |>
           map(fromJSON),
         message_len = map_int(message, \(m) {
           m$text |>
             replace_na("") |>
             str_c(collapse = " ") |>
             str_count("\\S+")
         }),
         context_len = map_int(message_history, \(m) {
           if (length(m) == 0) {
             return(0L)
           }
           m |>
             list_rbind() |>
             pull(text) |>
             replace_na("") |>
             str_c(collapse = " ") |>
             str_count("\\S+")
         }))
random_human_fixed <- human_selections_new |> 
  filter(condition == "random") |> 
  left_join(random_msgs |> 
              select(workerid, matcher_trialNum, matcher_repNum, 
                     message_length = message_len, context_length = context_len),
            by = join_by(workerid, matcher_trialNum, matcher_repNum)) |> 
  mutate(runId = as.character(workerid)) |>
  group_by(runId, condition, matcher_trialNum, matcher_repNum, 
           orig_trialNum, orig_repNum, gameId, target = correct_tangram) |>
  summarise(accuracy = mean(correct),
            message_length = mean(message_length),
            context_length = mean(context_length),
            .groups = "drop") |> 
  mutate(type = "human_naive")
```

```{r}
full_combined <- bind_rows(
  human_orig_fixed,
  yoked_human_fixed,
  shuffled_human_fixed,
  backward_human_fixed,
  random_human_fixed,
  logprobs_base
) |> 
  mutate(type = factor(type, levels = c("human_original", "human_naive", "model_qwen", "model_gemma", 
                                      "model_llama", "model_kimi")),
         condition = factor(condition, levels = c("yoked", "shuffled", "backward", "ablated", 
                                                  "other-within", "other-across", "random", "no context"))) |> 
  arrange(type, condition, runId, matcher_trialNum, feedback, image) |>
  group_by(type, condition, runId, target, feedback, image) |>
  mutate(matcher_repNum = ifelse(condition %in% c("shuffled", "random"), 0:5, matcher_repNum)) |> 
  ungroup()
```

```{r}
all_combined <- full_combined |> 
  rename(image_type = image) |> 
  pivot_wider(names_from = image_type,
              values_from = accuracy) |> 
  mutate(accuracy = image - `no image`)
```


# Plots
## Big plot
```{r eval=F}
p_matcher <- make_accuracy_plot(full_combined)
ggsave(here("figs", "matcher_plot.png"), p_matcher, width = 9, height = 4.5)
p_original <- full_combined |> 
  filter(condition %in% c("shuffled", "backward", "random", "no context")) |> 
  make_accuracy_plot(repnum_type = "original") +
  COL_SCALE_NOLIM
ggsave(here("figs", "original_plot.png"), p_original, width = 5.7, height = 4.5)
```

```{r}
p_big_matcher <- full_combined |> 
  filter(condition != "no context") |> 
  cast_random() |> 
  arrange(type, condition, gameId, runId, matcher_trialNum) |> 
  make_accuracy_plot() +
  theme(legend.position = "bottom")
p_big_orig <- full_combined |> 
  filter(condition %in% c("shuffled", "backward", "random", "no context")) |> 
  make_accuracy_plot(repnum_type = "original") +
  theme(legend.position = "none")
p_big <- (p_big_matcher | p_big_orig) + 
  plot_annotation(tag_levels = 'A') +
  plot_layout(
    widths = c(2.05, 1.05),
    guides = "collect") &
  theme(legend.position = "bottom")
ggsave(here("figs", "big_plot.png"), p_big, width = 10, height = 5.5)
```

## No image plots
```{r}
p_noimg_diff <- all_combined |> 
  make_accuracy_plot(ref_level = 0) +
  COL_SCALE_NOLIM +
  labs(y = "Accuracy difference (text-only − image–text)")
ggsave(here("figs", "diff_plot.png"), p_noimg_diff, width = 9, height = 4.5)
```


## Context length plot
```{r}
h_len <- full_combined |> 
  filter(type == "model_qwen") |> 
  ggplot(aes(x = context_length, fill = condition)) +
  geom_density(alpha = .2) +
  labs(x = "Context length", y = "Density", fill = "Condition")
p_len <- full_combined |> 
  filter(type == "model_qwen") |>
  ggplot(aes(x = context_length, y = accuracy, col = condition)) +
  geom_hline(yintercept = 1 / 12, lty = "dashed") +
  geom_point(position = position_jitter(width = .2), alpha = .05) +
  geom_smooth(method = "loess") +
  labs(x = "Context length", y = "Accuracy", col = "Condition")

p_full_len <- h_len / p_len +
  plot_layout(
    heights = c(1, 4),
    guides = "collect",
    axes = "collect") &
  theme(legend.position = "bottom") &
  guides(fill = "none")
```

## Model vs human plot
```{r}
yoked_data <- full_combined |> 
  filter(condition == "yoked",
         !str_starts(type, "human_")) |> 
  left_join(full_combined |> 
              filter(condition == "yoked",
                     type == "human_naive") |> 
              rename(orig_acc = accuracy) |> 
              select(-condition, -matcher_trialNum, -matcher_repNum, 
                     -message_length, -context_length, -type),
            by = join_by(gameId, orig_trialNum, orig_repNum, target)) |> 
  mutate(repnum = orig_repNum + 1,
         condition = "yoked")

shuffled_data <- full_combined |> 
  filter(condition == "shuffled",
         !str_starts(type, "human_")) |> 
  left_join(full_combined |> 
              filter(condition == "shuffled",
                     type == "human_naive") |> 
              rename(orig_acc = accuracy) |> 
              select(-condition, -message_length, -context_length, -type),
            by = join_by(gameId, runId, orig_trialNum, orig_repNum, 
                         matcher_trialNum, matcher_repNum, target)) |> 
  mutate(repnum = orig_repNum + 1) |> 
  group_by(gameId, orig_trialNum, orig_repNum, type) |>
  summarise(
    accuracy = mean(accuracy),
    orig_acc = mean(orig_acc),
    .groups = "drop"
  ) |> 
  mutate(condition = "shuffled")

backward_data <- full_combined |> 
  filter(condition == "backward",
         !str_starts(type, "human_")) |> 
  left_join(full_combined |> 
              filter(condition == "backward",
                     type == "human_naive") |> 
              rename(orig_acc = accuracy) |> 
              select(-condition, -matcher_trialNum, -matcher_repNum, 
                     -message_length, -context_length, -type),
            by = join_by(gameId, orig_trialNum, orig_repNum, target)) |> 
  mutate(repnum = orig_repNum + 1,
         condition = "backward")

random_data <- full_combined |> 
  filter(condition == "random",
         !str_starts(type, "human_")) |> 
  left_join(full_combined |> 
              filter(condition == "random",
                     type == "human_naive") |> 
              rename(orig_acc = accuracy) |> 
              select(-condition, -message_length, -context_length, -type),
            by = join_by(gameId, runId, orig_trialNum, orig_repNum, 
                         matcher_trialNum, matcher_repNum, target)) |> 
  mutate(repnum = orig_repNum + 1) |> 
  group_by(gameId, orig_trialNum, orig_repNum, type) |>
  summarise(
    accuracy = mean(accuracy),
    orig_acc = mean(orig_acc),
    .groups = "drop"
  ) |> 
  mutate(condition = "random")

comparison_data <- bind_rows(yoked_data, shuffled_data, backward_data, random_data) |> 
  mutate(type = fct_drop(type),
         condition = factor(condition, levels = c("yoked", "shuffled", "backward", "random")))
```

```{r}
comparison_plot <- comparison_data |> 
  ggplot(aes(x = orig_acc, y = accuracy, col = type)) +
  geom_abline(slope = 1, intercept = 0, lty = "dashed") +
  geom_point(alpha = .1) +
  geom_smooth(method = "lm") +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) +
  facet_grid(~condition,
             labeller = as_labeller(str_to_sentence)) +
  COL_SCALE_NOLIM +
  labs(
    x = "Naïve human accuracy",
    y = "Model accuracy",
    col = "Matcher"
  ) +
  theme(legend.position = "bottom",
        legend.box.spacing = unit(0.5, "pt"))

ggsave(here("figs", "comparison_plot.png"), comparison_plot, width = 10.5, height = 3.2)
```

```{r}
comparison_cors <- comparison_data |> 
  group_by(type, condition, gameId, orig_trialNum) |> 
  summarise(
    accuracy = mean(accuracy),
    orig_acc = mean(orig_acc),
    .groups = "drop"
  ) |>
  group_by(type, condition) |> 
  summarise(
    correlation = cor(orig_acc, accuracy, use = "pairwise.complete.obs"),
    spearman = cor(orig_acc, accuracy, method = "spearman", use = "pairwise.complete.obs"),
    .groups = "drop"
  )

cor_text <- comparison_cors |> 
  group_by(condition) |> 
  summarise(
    cor_text = glue(
    'plain(\"model \") * italic(r) %in% \"[{min(correlation) |> strip_zero()}, ',
    '{max(correlation) |> strip_zero()}]\"'),
  )
```

Human split-half
```{r}
all_human <- bind_rows(human_selections, human_selections_new |> 
                         mutate(condition = ifelse(condition == "backwards", "backward", condition)))
workers <- all_human |> 
  distinct(workerid)

sh_cors <- map(1:1000, \(i) {
  set.seed(i)
  
  workers_split <- workers |> 
    mutate(split = sample(1:2, n(), replace = TRUE))
  
  sh_split <- all_human |> 
    left_join(workers_split, by = "workerid") |>
    group_by(condition, split, gameId, orig_trialNum, orig_repNum, 
             correct_tangram) |>
    summarise(accuracy = mean(correct),
              .groups = "drop") |>
    pivot_wider(names_from = split, values_from = accuracy, names_prefix = "split_") |>
    filter(!is.na(split_1) & !is.na(split_2)) |>
    group_by(condition) |> 
    summarise(
      cor = cor(split_1, split_2, use = "pairwise.complete.obs"),
      spearman = cor(split_1, split_2, method = "spearman", use = "pairwise.complete.obs"),
      .groups = "drop"
    ) |> 
    mutate(split = i)
  sh_split
}) |> list_rbind()

sh_cors_sum <- sh_cors |> 
  group_by(condition) |> 
  summarise(
    mean_cor = mean(cor),
    sd_cor = sd(cor),
    median_cor = median(cor),
    ci_lower = quantile(cor, .025),
    ci_upper = quantile(cor, .975),
    .groups = "drop"
  )

sh_sum_text <- sh_cors_sum |> 
  mutate(
    sh_text = glue(
      'plain(\"human \") * italic(r) == \"{mean_cor |> strip_zero()} [{ci_lower |> strip_zero()}, ',
    '{ci_upper |> strip_zero()}]\"')
  ) |> 
  select(condition, sh_text)
```

```{r}
full_cor_text <- cor_text |> 
  left_join(sh_sum_text, by = join_by(condition)) |> 
  mutate(cor_text = glue("atop({cor_text}, {sh_text})")) |> 
  mutate(condition = factor(condition, 
                           levels = c("yoked", "shuffled", "backward")))
```


Plot
```{r}
comparison_plot_text <- comparison_plot +
  geom_label(
    aes(x = 1, y = 0, label = cor_text),
    col = "black",
    data = full_cor_text,
    parse = TRUE,
    size = 3,
    hjust = 1,
    vjust = 0
  )
ggsave(here("figs", "comparison_plot.png"), comparison_plot_text, width = 9, height = 3)
```


# Modelling
```{r}
model_data <- full_combined |> 
  filter(type != "human_original") |> 
  left_join(full_combined |> 
              filter(type == "human_original") |> 
              rename(orig_acc = accuracy) |> 
              select(-condition, -matcher_trialNum, -matcher_repNum, 
                     -message_length, -context_length, -type),
            by = join_by(gameId, orig_trialNum, orig_repNum, target)) |> 
  mutate(
    type = fct_drop(type),
    across(c(type, condition), \(x) x |> `contrasts<-`(value = contr.sum(nlevels(x))))
  )

model <- lmer(
  accuracy ~ type * condition * (poly(matcher_repNum, 5) + poly(orig_repNum, 5)) + 
    type * (poly(message_length, 2) + poly(context_length, 2)) + orig_acc + (1 | gameId) + (1 | target),
  data = model_data,
  # prior = c(
  #   set_prior("normal(0, 1)", class = "b"),
  #   set_prior("normal(0, 1)", class = "Intercept")
  # ),
  # iter = 4000, warmup = 2000, chains = 4, cores = 4
)

summary(model)
```

```{r}
library(emmeans)
emt <- emtrends(model, ~ type, var = "context_length", max.degree = 2)
emt_pairs <- pairs(emt)
```

```{r, eval=F}
# doesn't work---lots of nonEst in the emm
emm <- emmeans(model, ~ type * matcher_repNum | condition , 
               at = list(matcher_repNum = 1:6))
cons <- contrast(emm, "pairwise", by = c("type", "matcher_repNum"),
                 adjust = "tukey")
```

# Item-level analysis
```{r}
p_items <- full_combined |> 
  mutate(target = fct_reorder(target, accuracy, .na_rm = TRUE)) |>
  ggplot(aes(x = target, y = accuracy, col = type)) +
  geom_hline(yintercept = 1 / 12, lty = "dashed") +
  stat_summary(
      aes(group = interaction(gameId, type, target)),
      fun = mean, geom = "point", alpha = .2,
      position = position_jitter(width = .2)
    ) +
  stat_summary(fun.data = mean_cl_boot, 
               geom = "pointrange", size = .4,
               position = position_dodge(width = .2)) +
  facet_wrap(~condition, nrow = 2,
             labeller = as_labeller(str_to_sentence)) +
  COL_SCALE +
  labs(x = "Tangram", y = "Accuracy", col = "Matcher") +
  theme(legend.position = "bottom")
ggsave(here("figs", "items_plot.png"), p_items, width = 9, height = 5.5)
```

# Learning
```{r}
accs_full <- map(logprobs_full_base, calculate_accuracies)

orig_acc <- human_orig |> 
  mutate(runId = as.character(1),
         orig_accuracy = accuracy,
         accuracy = as.numeric(accuracy > 0),
         target_repNum = matcher_repNum) |> 
  calculate_accuracies(join_only = TRUE) |> 
  mutate(accuracy = orig_accuracy) |> 
  select(-orig_accuracy)

naive_acc <- human_selections |> 
  bind_rows(human_selections_new) |> 
  mutate(type = "human_naive",
         runId = as.character(workerid), 
         accuracy = as.numeric(correct),
         condition = ifelse(condition == "backwards", "backward", condition)) |> 
  cast_random() |> 
  rename(target = correct_tangram) |> 
  select(type, condition, gameId, runId, orig_trialNum, orig_repNum,
         matcher_trialNum, matcher_repNum, target, accuracy) |> 
  arrange(type, condition, gameId, runId, matcher_trialNum) |>
  group_by(type, condition, gameId, runId, target) |>
  mutate(target_repNum = ifelse(condition %in% c("shuffled", "random"), 0:5, matcher_repNum)) |> 
  calculate_accuracies(join_only = TRUE)

all_acc <- bind_rows(
  bind_rows(accs_full), orig_acc, naive_acc
) |> 
  mutate(
    type = factor(type, 
                  levels = c("human_original", "human_naive", "model_qwen", "model_gemma", 
                             "model_llama", "model_kimi")),
    condition = factor(condition, 
                       levels = c("yoked", "shuffled", "backward", "ablated", 
                                  "other-within", "other-across", "random", "no context"))
  )
```


```{r}
p_learning <- all_acc |> 
  filter(!is.na(prev_accuracy),
         condition %in% c("yoked", "shuffled", "backward", "random")) |> 
  mutate(target_repNum = target_repNum + 1,
         prev_accuracy = ifelse(prev_accuracy == 1, 
                                "Prev correct", "Prev incorrect")) |> 
  ggplot(aes(x = target_repNum, y = accuracy, 
             col = type)) +
  geom_hline(yintercept = 1 / 12, lty = "dashed") +
  stat_summary(
      aes(group = interaction(gameId, type)),
      fun = mean, geom = "line", alpha = .15,
      show.legend = TRUE
    ) +
  stat_smooth(
    method = "loess", formula = y ~ x, se = FALSE,
    span = .9,
    show.legend = TRUE
  ) +
  stat_summary(
    fun.data = mean_cl_boot, geom = "pointrange",
    lty = "solid",
    show.legend = TRUE
  ) +
  facet_grid(prev_accuracy ~ condition, # nrow = 2,
             labeller = as_labeller(str_to_sentence)) +
  scale_x_continuous(breaks = 2:6) +
  # scale_shape_manual(values = c(1, 16)) +
  # scale_linetype_manual(values = c("longdash", "solid")) +
  COL_SCALE +
  labs(
    x = glue("Matcher repetition number"),
    y = "Accuracy", col = "Matcher"# ,
    # lty = "Previous accuracy", shape = "Previous accuracy"
  )
ggsave(here("figs", "learning_plot.png"), p_learning, width = 9, height = 4.5)
```

# Confusion matrices
Do confusion matrices, and look at RSA across conditions
```{r}
mat_conf <- bind_rows(logprobs_full_base) |> 
  group_by(type, condition, target, selection) |>
  summarise(
    prob = mean(prob, na.rm = TRUE),
    .groups = "drop"
  )
```

```{r}
mat_conf |> 
  filter(type == "model_qwen", condition == "yoked") |> 
  ggplot(aes(x = target, y = selection, fill = prob)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "#7212b3", limits = c(0, 1)) +
  labs(x = "Target", y = "Selection", fill = "Probability")
```

```{r}
mat_conf_wide <- mat_conf |> 
  pivot_wider(names_from = condition,
              values_from = prob)
mat_conf_rsa <- mat_conf_wide |> 
  select(-target, -selection) |> 
  nest(data = -type) |> 
  mutate(correl = map(data, \(d) cor(d |> as.matrix()) |> list()))
```

```{r}
mat_conf_rsa_long <- mat_conf_rsa |> 
  select(-data) |> 
  mutate(correl = map(correl, \(c) as_tibble(c[[1]], rownames = "condition1"))) |> 
  unnest(correl) |> 
  pivot_longer(cols = yoked:`no context`,
               names_to = "condition2",
               values_to = "cor") |> 
  mutate(across(starts_with("condition"), 
                \(c) c |> str_to_sentence() |> 
                  factor(levels = c("Yoked", "Shuffled", "Backward", "Ablated", 
                                    "Other-within", "Other-across", "Random", "No context"))))
```

```{r}
p_rsa <- mat_conf_rsa_long |> 
  mutate(type = case_when(
           type == "model_gemma" ~ "Gemma 3",
           type == "model_qwen" ~ "Qwen 2.5 VL",
           type == "model_llama" ~ "Llama 3.2",
           type == "model_kimi" ~ "Kimi VL"
         ) |> factor(levels = c("Qwen 2.5 VL", "Gemma 3", "Llama 3.2", "Kimi VL"))) |>
  ggplot(aes(x = condition2, y = condition1, fill = cor)) +
  geom_tile() +
  facet_grid(~type) +
  scale_fill_gradient(low = "white", high = "#7212b3", limits = c(0, 1)) +
  labs(x = "", y = "", fill = "Correlation") +
  theme(axis.text.x = element_text(hjust = 1, angle = 45))
ggsave(here("figs", "rsa_plot.png"), p_rsa, width = 10, height = 3.3)
```

# Error analysis
```{r}
# reload texts
game_texts <- original_chat |>
  mutate(
    role = case_when(
      role == "speaker" ~ "describer",
      role == "listener" ~ "matcher"
    )) |> 
  rename(orig_trialNum = trialNum,
         orig_repNum = repNum) |> 
  select(gameId, orig_trialNum, orig_repNum, role, text, condition) |> 
  mutate(role_text = glue("{role}: {text}")) |>
  group_by(gameId, orig_trialNum, orig_repNum, condition) |> 
  summarise(role_text = str_c(role_text, collapse = "\n"),
            text = str_c(text, collapse = " "),
            .groups = "drop")

# lemmatise
ud_model <- udpipe_load_model(here("resources", "english-ewt-ud-2.5-191206.udpipe"))
game_texts_lemmas <- game_texts |> 
  mutate(ud = map(text, \(t) udpipe_annotate(ud_model, x = str_to_lower(t)) |> 
                    as_tibble() |> select(lemma) |> 
                    # manual corrections
                    mutate(lemma = case_when(
                      lemma == "triangles" ~ "triangle", 
                      lemma == "ears" ~ "ear",
                      lemma == "backwards" ~ "backward",
                      lemma == "forwards" ~ "forward",
                      lemma == "kneeling" ~ "kneel",
                      lemma == "leaning" ~ "lean",
                      lemma == "shapes" ~ "shape",
                      lemma == "shoulders" ~ "shoulder",
                      lemma == "sitting" ~ "sit",
                      lemma == "towards" ~ "toward",
                      lemma == "upwards" ~ "upward",
                      lemma == "walking" ~ "walk",
                      .default = lemma)) |> 
                    distinct())) |> 
  unnest(ud)
```

```{r}
random_combined <- full_combined |> 
  filter(condition == "random") |> 
  group_by(type, gameId, orig_trialNum, orig_repNum, target) |>
  summarise(accuracy = mean(accuracy),
            .groups = "drop")
```

```{r}
# error_analysis <- random_combined |> 
#   pivot_wider(names_from = type, values_from = accuracy) |> 
#   mutate(across(starts_with("model_"), \(x) x - human_naive, .names = "{.col}_vs_naive"))

error_analysis <- random_combined |> 
  mutate(type = ifelse(str_starts(type, "model_"), "model", "human_naive")) |> 
  group_by(type, gameId, orig_trialNum, orig_repNum, target) |>
  summarise(accuracy = mean(accuracy),
            .groups = "drop") |> 
  pivot_wider(names_from = type, values_from = accuracy) |> 
  mutate(diff = model - human_naive) |> 
  left_join(game_texts, by = join_by(gameId, orig_trialNum, orig_repNum))
```

```{r}
error_analysis_lemmas <- error_analysis |> 
  left_join(game_texts_lemmas, by = join_by(gameId, orig_trialNum, orig_repNum)) |> 
  group_by(lemma) |> 
  summarise(
    n = n(),
    n_games = n_distinct(gameId),
    mean_diff = mean(diff, na.rm = TRUE),
    sd_diff = sd(diff, na.rm = TRUE),
    se_diff = sd_diff / sqrt(n),
    ci_lower = mean_diff - qt(0.975, n - 1) * se_diff,
    ci_upper = mean_diff + qt(0.975, n - 1) * se_diff,
    .groups = "drop"
  ) |> 
  filter(n > 1) |> 
  mutate(diff_vs_grand_mean = mean_diff - mean(error_analysis$diff, na.rm = TRUE),
         ci_lower_vs_grand_mean = ci_lower - mean(error_analysis$diff, na.rm = TRUE),
         ci_upper_vs_grand_mean = ci_upper - mean(error_analysis$diff, na.rm = TRUE),
         ci_vs_grand_mean = case_when(
           ci_lower > mean(error_analysis$diff, na.rm = TRUE) ~ "above",
           ci_upper < mean(error_analysis$diff, na.rm = TRUE) ~ "below",
           TRUE ~ "includes"
         ))

error_analysis_lemmas |> 
  filter(n > 10) |> 
  write_csv("mean_diff_by_lemma.csv")
```


```{r}
# quick n dirty plot
# body parts are bad: arm, leg, foot, hand, head
# directions are bad: up, left, right, but not down
# shapes ns: diamond, square, triangle, rectangle

# good words: bunny, candle (but just two games)

SEARCH_WORD = "candle"

error_analysis |> 
  mutate(!!SEARCH_WORD := str_detect(text, SEARCH_WORD)) |>
  ggplot(aes(x = diff, fill = !!sym(SEARCH_WORD))) +
  geom_density(alpha = .6)

# t-test for search vs full distribution
t.test(
  error_analysis |> pull(diff),
  error_analysis |> filter(str_detect(text, SEARCH_WORD)) |> pull(diff)
)
```

Trying a different plot
```{r}
lemmas_annotated <- read_csv(here("resources", "lemmas_annotated.csv"))

error_analysis_lemmas_annotated <- error_analysis_lemmas |> 
  left_join(lemmas_annotated, by = join_by(lemma)) |> 
  filter(!is.na(category)) |> 
  mutate(category = fct_reorder(category, diff_vs_grand_mean),
         lemma = fct_reorder(lemma, diff_vs_grand_mean, .desc = TRUE))

p_errors <- error_analysis_lemmas_annotated |> 
  mutate(signif = ifelse(ci_vs_grand_mean == "includes", "no", "yes")) |>
  ggplot(aes(x = diff_vs_grand_mean, y = lemma, col = category)) +
  geom_vline(xintercept = 0, lty = "dashed") +
  geom_errorbarh(aes(xmin = ci_lower_vs_grand_mean, xmax = ci_upper_vs_grand_mean),
                 height = .4) +
  geom_point(aes(shape = signif), size = 2) +
  facet_grid(category ~ ., scales = "free_y", space = "free_y") +
  scale_shape_manual(values = c(1, 16)) +
  labs(x = "Centred discrepancy (model – human)", y = "Lemma") +
  theme(legend.position = "none")

ggsave(here("figs", "error_plot.png"), p_errors, width = 6, height = 8)
```

For completeness
```{r}
error_analysis_lemmas |> 
  left_join(lemmas_annotated, by = join_by(lemma)) |> 
  filter(is.na(category), n >= 20, n_games >= 4) |> 
  mutate(lemma = fct_reorder(lemma, diff_vs_grand_mean, .desc = TRUE)) |> 
  mutate(signif = ifelse(ci_vs_grand_mean == "includes", "no", "yes")) |>
  ggplot(aes(x = diff_vs_grand_mean, y = lemma, col = category)) +
  geom_vline(xintercept = 0, lty = "dashed") +
  geom_errorbarh(aes(xmin = ci_lower_vs_grand_mean, xmax = ci_upper_vs_grand_mean),
                 height = .4) +
  geom_point(aes(shape = signif), size = 2) +
  facet_grid(category ~ ., scales = "free_y", space = "free_y") +
  scale_shape_manual(values = c(1, 16)) +
  labs(x = "Centred discrepancy (model – human)", y = "Lemma") +
  theme(legend.position = "none")
```

## Span labelling
```{r}
spans <- read_csv(here("data", "original_chat_spans.csv"))

spans_cleaned <- spans |> 
  select(-raw_spans) |> 
  mutate(n_literal = str_count(spans, "\"literal\""),
         n_metaphorical = str_count(spans, "\"metaphorical\""),
         n_partbased = str_count(spans, "\"part_based\""),
         n_holistic = str_count(spans, "\"holistic\""),
         n_spans = str_count(spans, "\"span\"")) |> 
  group_by(gameId, trialNum, repNum) |> 
  summarise(across(starts_with("n_"), sum),
            .groups = "drop") |> 
  mutate(prop_metaphorical = n_metaphorical / n_spans,
         prop_holistic = n_holistic / n_spans)
```

```{r}
library(irr)
spans_irr <- read_csv(here("data", "original_chat_spans_irr.csv"))

spans_irr_cleaned <- spans_irr |> 
  select(-raw_spans) |> 
  mutate(n_literal = str_count(spans, "\"literal\""),
         n_metaphorical = str_count(spans, "\"metaphorical\""),
         n_partbased = str_count(spans, "\"part_based\""),
         n_holistic = str_count(spans, "\"holistic\""),
         n_spans = str_count(spans, "\"span\"")) |> 
  group_by(gameId, trialNum, repNum) |> 
  summarise(across(starts_with("n_"), sum),
            .groups = "drop") |> 
  mutate(prop_metaphorical = n_metaphorical / n_spans,
         prop_holistic = n_holistic / n_spans) |> 
  select(gameId, trialNum, repNum, 
         prop_metaphorical_recode = prop_metaphorical, 
         prop_holistic_recode = prop_holistic) |> 
  left_join(spans_cleaned |> 
              select(gameId, trialNum, repNum, 
                     prop_metaphorical_orig = prop_metaphorical, 
                     prop_holistic_orig = prop_holistic),
            by = join_by(gameId, trialNum, repNum))

icc(spans_irr_cleaned |> select(prop_metaphorical_orig, prop_metaphorical_recode),
    model = "twoway", type = "agreement", unit = "single")
icc(spans_irr_cleaned |> select(prop_holistic_orig, prop_holistic_recode),
    model = "twoway", type = "agreement", unit = "single")
```


Humans
```{r}
random_human_spans <- random_human |> 
  left_join(spans_cleaned, 
            by = join_by(gameId, orig_trialNum == trialNum, 
                         orig_repNum == repNum))

p_hum_met <- random_human_spans |> 
  ggplot(aes(x = prop_metaphorical, y = accuracy)) +
  geom_point(alpha = .1, position = position_jitter(height = .1, width = .05)) +
  geom_smooth(method = "glm", method.args = list(family = "quasibinomial")) +
  labs(x = "Proportion of metaphorical spans", y = "Accuracy")

p_hum_hol <- random_human_spans |> 
  ggplot(aes(x = prop_holistic, y = accuracy)) +
  geom_point(alpha = .1, position = position_jitter(height = .1, width = .05)) +
  geom_smooth(method = "glm", method.args = list(family = "quasibinomial")) +
  labs(x = "Proportion of holistic spans", y = "Accuracy")

p_hum_met + p_hum_hol
```

```{r}
spans_cleaned |> 
  ggplot(aes(x = repNum, y = prop_metaphorical)) +
  geom_point(alpha = .5, position = position_jitter(width = .2, height = .2)) +
  geom_smooth(method = "lm") +
  labs(x = "Repetition number", y = "Proportion of metaphorical spans")
```

```{r}
spans_cleaned |> 
  ggplot(aes(x = repNum, y = prop_holistic)) +
  geom_point(alpha = .5, position = position_jitter(width = .2, height = .2)) +
  geom_smooth(method = "lm") +
  labs(x = "Repetition number", y = "Proportion of holistic spans")
```

```{r}
spans_cleaned |> 
  ggplot(aes(x = prop_metaphorical, y = prop_holistic)) +
  geom_point(alpha = .5) +
  geom_smooth(method = "lm") +
  labs(x = "Proportion of metaphorical spans", y = "Proportion of holistic spans")
```


